---
title: "Midpoint Deliverable Group 3"
author: "Russell Eng, Anne Lin, Xinmeng Zhang, Neil Sun"
date: "08/13/2021"
output:
  prettydoc::html_pretty: default
  html_document: default
---

```{r setup, include = F, message = F}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
library("sf")
library("knitr")
library("kableExtra")
df <- read.csv("data/us_state_vaccinations.csv")
states <- read_sf("data/us_map/tl_2020_us_state.shp")
sex_df <- read.csv("data/sex_of_people_fully_vaccinated.csv", header = TRUE, 
  strip.white = TRUE, sep = ",")
population <- read.csv("data/population.csv")
blue_red_states <-
  read.csv("data/democratic_vs_republican_votes_by_usa_state_2020.csv")
source("scripts/summary info table.R")
source("scripts/chart 1.R")
source("scripts/chart 2.R")
source("scripts/chart 3.R")
source("scripts/summary info.R")
options(scipen = 999)

```

## Introduction of Problem Domain and Dataset
***
This project is about the **vaccinations of COVID-19 vaccines** in the US. 
Given that the Delta variant is causing a new surge of COVID-19 cases, it is key 
that all people get their vaccination as soon as possible. We are eager to know 
_how people of different states react to vaccination_. By breaking down the 
vaccinated population by state, it would be easier to create targeting 
incentives, promoting vaccination to fight against the more transmissible Delta 
variant. Our data set is downloaded from the [CDC website](https://covid.cdc.gov/covid-data-tracker/#vaccination-trends) and 
contains information including the amount of vaccines distributed to and 
administered in each state over the course of 228 days, 
from 12/20/2020 to 08/04/2021. 


## Summary Information
***
For the summary, we included the following information:

- **The column names, including:**

>`r summary_info(df, blue_red_states)$col_names`

- **the total number of vaccinations nationwide as of:**

>`r summary_info(df, blue_red_states)$total_vacc_0804`

- **the state with the highest percentage of fully vaccinated population as of 08/04/2021:**

>`r summary_info(df, blue_red_states)$max_fully_rate_state`

- **the state with the highest number of vaccination distribution as of 08/04/2021:**

>`r summary_info(df, blue_red_states)$most_distributed_state`

- **the number of people fully vaccinated in Washington state as of 08/04/2021:**

>`r summary_info(df, blue_red_states)$wa_fully_vacc_0804`

## Aggregate Table
***
Below is a table that summarizes the number of average daily vaccinations and 
what day in the week has the most vaccinations in each state.

```{r summary_table, echo = F, message = F}
kbl(get_summary_table(df, population),
    col.names = c("Location", "Political Stance", "Most Vaccination Weekday",
                "Daily Average Vaccination",
                "Percentage of Daily Vaccinated Population",
                "Fully-Vaccinated Population",
                "Total Vaccinations")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

We sorted the table in descending order based on the
_Percentage of Daily Vaccinated Population_, and Vermont tuned out to be the
state with the most percentage of population vaccinated every day. This result
resonates well with one of the conclusion we drew from the summary: Vermont has
the highest fully-vaccinated population ratio among all states.

We were also interested in knowing on what day in the week people are the most
willing to go get vaccinated in the week and concluded from that **Saturday**
has the most vaccinations; this is likely because many people work on weekdays and
are not able to go get their vaccination on such days. Another thing we found
interesting from the table is that the states with higher rates of fully
vaccinated population are **democratic states**. 


## Leaflet Map
***
Below is a map that shows the percentage of the population that is
fully vaccinated in each state.

```{r chart 1, echo=FALSE, message = FALSE, warning = FALSE} 
map_chart(df, states)
```

We used _Leaflet_, the most common open source for interactive map charts, to
generate the map and facilitate further web-page programming using shiny().
While there are other districts, departments, and entities listed in the data,
we only selected the 51 states/districts in the US as we hoped to see the
distribution of the fully vaccinated population in each states. Based on the
map, we are able to conclude that **states on the west coast and the upper east
coast are the most vaccinated states**, while many of the southern states are
the least vaccinated states. This result can be tied back to the aggregated
table and illustrate the fact that democratic states tend to have the highest
rate of vaccinated population.

## Area Graph
***
Below is a area graph that shows the trend of vaccinations in weeks starting
from week 1 in 2021 to week 31 in 2021 (01/04/2012 - 08/04/2021).
*Note that week 31 only has data for three days.

```{r chart 2, echo = F, message = FALSE} 
line_chart(df)
```

We used an area chart for our data set with a great scale of quantitative data,
as it is one of the best way for communicating a trend over the course of a
specified period. 

The area graph shows that the weekly vaccination in the US in 2021 started at
around 4 million at the beginning and **increased until it peaked in week 15**,
which is mid-April. However, the number of weekly vaccinations started to
**decrease afterwards**. The reason why this sudden and sharp decline is yet to
be discussed and will be included in the final submission.

## Bar Chart 
***
Below is a bar chart that shows the percentage of males and females
in the US population, the vaccinated population, and the people vaccinated in
the last 14 days.

```{r chart 3, echo = F, message=FALSE} 
bar_chart(sex_df)
```

Bar chart is a useful layout to compare the differences between genders in our
data set. According the the three bars shown above, while the percentage of
females and males in the US are nearly equal, more females are vaccinated than
males in the fully-vaccinated population. At this point in time, we can conclude
that **females are generally more open to vaccination than males in the US**.
